Geometric variational problems

### 2020-11-09 NSF2020/Description/specificaim2_ver2.tex
Update specificaim2_ver2.tex
Still working at SA2

%\subsection{Specific Aim 2: Functional Materials}
\subsection{Specific Aim 2: Hydrodynamics}
\label{subsec:specific_aim_2}
Specific Aim 1 dealt with the collective material property of amphiphiles which does not involve the movement  of the solvent.
In Specific Aim 2, we include hydrodynamics effects. Hydrodynamic effects are relevant
because the rates of biological functions like fusion, fission, and pore dynamics rely on viscous dissipation \cite{RYHAM20112929},
and can determine which local equilibria the system tends towards. 
Additionally, coarse-grained and molecular dynamics theories include water either explicitly or implicitly in their models.

A further area of relevance is in the fabrication of complex microscopic three-dimensional structures \cite{Cho2010}.
Over the past decade, there has been an explosion of interest in small-scale processes that utilize capillary forces dominate,  van der Waals interactions and thermal noise  \cite{Zhang2017},
to coordinated movement and bind  material subunits. Two prominent fabrication techniques are capillary origami 
\cite{Pandey2011,Leong2007,Reynolds2019} and colloidal self-assembly \cite{Dasgupta2017,Siontorou2017}. Capillary origami uses principles of elastocapilirity wherein elastic solids deform under surface tension \cite{Bico2018,VanHonschoten2010}. In self-assembly of colloids, 
particles exhibit similar thermodynamic behavior of molecular systems, except that the behavior occurs over observable, long time scales \cite{Zhang2017}.

The hydrodynamic interactions enter through the transmission of viscous stresses from one particle to the another.
This involves solving the system

\setcounter{midequation}{\theequation}
\addtocounter{midequation}{2}

@@

Numerically solving \eqref{St1}--\eqref{HSB2} is a nontrivial matter.
First, we have to carefully map the particle configurations because the domain $\Omega(t)$ and boundary conditions $f_i$ are constantly changing.
The particle dynamics in \eqref{Hupdate1} and in \eqref{SGD} both involve solving
screened Laplace boundary value problems on evolving domains, but differ in that \eqref{Hupdate1} has an additional Stokes flow
generating the dissipation. When there is no background flow ${\bf u}_{\infty} = {\bf 0}$, the system obeys the energy law
\begin{equation}
\frac{\dif}{\dif t} \Phi + \int_{\Omega(t)} \mu |\nabla {\bf u} + \nabla {\bf u}^T|^2 \,dx = 0
\end{equation}
and the particle configurations tend toward a local equilibrium.

A nonzero background flow continually injects energy into the system. We deal with the background flow using
the Power and Miranda representation 
\begin{equation}
\label{PowerMiranda}
{\bf u} = {\bf u}_{\infty} + D\eta + \sum_{i=1}^N S(\cdot, {\bf a}_i) {\bf F}_i + R(\cdot, {\bf a}_i) {\bf G}_i.
\end{equation}
where $S(\cdot, {\bf a}_i)$ and $ R(\cdot, {\bf a}_i)$ are stokeslets and rotlets supported at the particle centers ${\bf a}_i.$
The symbol $D$ denotes the Stokes double layer integral operator (see Specific Aim 3).   
The representation \eqref{PowerMiranda} automatically satisfies \eqref{St1}, \eqref{St2}, \eqref{St4} and \eqref{HSB1}.
We solve for the unknown density $\eta : \partial \Omega \to \mathbb{R}$ by requiring that jump in viscous stress
vanished across $\partial P_i$. This jump condition guarantees that ${\bf u}$ is a rigid motion within each particle,
and by continuity, equals a rigid motion on the particle boundary as required by \eqref{St3}.

We prove a, what to our knowledge is new, result concerning the evaluation of Maxwell-type stresses. 
\begin{proposition}
  \label{prop:recip}
  Suppose that $u = \sum_{i=1}^N u_i$ where $u_i$ is a solution of the screened Laplace equation
  in $\mathbb{R}^3 \setminus P_i$ for $i=1,\dots, N.$ Then 
\begin{equation}
{\bf F}_i = \sum_{j \neq i} \int_{\partial P_i}[\boldsymbol{\sigma}_{ij} + \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S,\quad
{\bf G}_i = \sum_{j \neq i} \int_{\partial P_i} {\bf x} \times [\boldsymbol{\sigma}_{ij} + \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S, \quad i = 1\dots, N.
\end{equation}
where $\boldsymbol{\sigma}_{ij} = \rho^{-1} u_iu_j {\bf I} + \rho(\nabla u_i \cdot \nabla u_j {\bf I} - 2 \nabla u_i \otimes \nabla u_j)$ and
$[\cdot]$ denotes the jump in stress across $\partial \Omega.$ 
\end{proposition}
As described in Specific Aim 3, we solve \eqref{SL1}--\eqref{SL3}  using a double layer potential representation.
While accurate in the far-field, this representation involves a strongly singular integral which
leads to large numerical errors when evaluating the the hydrophobic stress \eqref{stress} along the particle boundaries.
The identity in Proposition \ref{prop:recip} gets around this critical problem by utilizing the fact that $u_j$, $j\neq i$ is smooth and the normal derivative
of  $u_i$ continuous across $\partial P_i$.

>>>>

%\subsection{Specific Aim 2: Functional Materials}
\subsection{Specific Aim 2: Hydrodynamics}
\label{subsec:specific_aim_2}
Specific Aim 1 dealt with the collective material property of amphiphiles which does not involve the movement of the solvent.
In Specific Aim 2, we include hydrodynamics effects by embedding the particles in solvent. 
The fluid is modeled by the Stokes equations an incompressible, zero Reynolds-number fluid, and coupled to the screened-Laplace equation \eqref{SL}
through the viscous and hydrophobic stress balance. We develope a numerical solver for the mobility problem. 
This leads to simulations of bilayers in external flows where the bilayers are represented by a self-assebled collection of rigid particles. 
Finally, we work toward three-dimensional solvers with application in the capillary driven fabrication of micro-materials.

Hydrodynamic effects are relevant detail 
because the rates of biological functions like fusion, fission, and pore dynamics rely on viscous dissipation \cite{RYHAM20112929}. 
The hydrophobic attraction still causes the particles to self-assemble, but the trajectories and even configurations are changed
by the fact that the particles must now move through an ambient environment. 
Additionally, coarse-grained and molecular dynamics theories include water either explicitly or implicitly in their models.

The hydrodynamic interactions enter through the transmission of viscous stresses from one particle to the another.  
We must solve the system 
\setcounter{midequation}{\theequation}
\addtocounter{midequation}{2}

@@

Numerically solving \eqref{St1}--\eqref{HSB2} is a nontrivial matter.
First, we have to carefully map the particle configurations because the domain $\Omega(t)$ and boundary conditions $f_i$ are constantly changing.
The particle dynamics in \eqref{Hupdate1} and in \eqref{SGD} both involve solving
screened Laplace boundary value problems on evolving domains, but differ in that \eqref{Hupdate1} has an additional Stokes flow
generating the dissipation. When there is no background flow ${\bf u}_{\infty} = {\bf 0}$, the system obeys the energy law
$\frac{\dif}{\dif t} \Phi + \int_{\Omega(t)} \mu |\nabla {\bf u} + \nabla {\bf u}^T|^2 \,dx = 0,$ 
and the particle configurations tend toward a local equilibrium.

We use \eqref{St1}--\eqref{HSB2} to simulate two-dimensional vesicle bilayer in external flows. 
The vesicle bialyer is formed by a many-body  collection of amphiphilic particles. The inner, bilayer leaflet consists
of particles pointing their hydrophobic tails away from the vesicle center; the tails of the outer leaflet point inward.
We consider two kinds of external flows: shear flow and extensional flow. 
Specifically, we deal with the background flow using the representation 
\begin{equation}
\label{PowerMiranda}
{\bf u} = {\bf u}_{\infty} + D\eta + \sum_{i=1}^N S(\cdot, {\bf a}_i) {\bf F}_i + R(\cdot, {\bf a}_i) {\bf G}_i.
\end{equation}
where $S(\cdot, {\bf a}_i)$ and $ R(\cdot, {\bf a}_i)$ are stokeslets and rotlets supported at the particle centers ${\bf a}_i.$
The symbol $D$ denotes the Stokes double layer integral operator (see Specific Aim 3).   
The representation \eqref{PowerMiranda} automatically satisfies \eqref{St1}, \eqref{St2}, \eqref{St4} and \eqref{HSB1}.
We solve for the unknown density $\eta : \partial \Omega \to \mathbb{R}$ by requiring that jump in viscous stress
vanished across $\partial P_i$. This jump condition guarantees that ${\bf u}$ is a rigid motion within each particle,
and by continuity, equals a rigid motion on the particle boundary as required by \eqref{St3}.

The background flows allow us to study further properties of amphiphilic particle collections. 
Area conservation
Volumne conservation
Analyzing the dynamics compared to previous works
Interaction between the fluid and membrane

We prove a, to our knowledge new, result concerning the evaluation of Maxwell-type stresses. 
\begin{proposition}
  \label{prop:recip}
  Suppose that $u = \sum_{i=1}^N u_i$ where $u_i$ is a solution of the screened Laplace equation
  in $\mathbb{R}^3 \setminus P_i$ for $i=1,\dots, N.$ Then 
\begin{equation}
{\bf F}_i = \sum_{j \neq i} \int_{\partial P_i}[\boldsymbol{\sigma}_{ij} + \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S,\quad
{\bf G}_i = \sum_{j \neq i} \int_{\partial P_i} {\bf x} \times [\boldsymbol{\sigma}_{ij} + \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S, \quad i = 1\dots, N.
\end{equation}
where $\boldsymbol{\sigma}_{ij} = \rho^{-1} u_iu_j {\bf I} + \rho(\nabla u_i \cdot \nabla u_j {\bf I} - 2 \nabla u_i \otimes \nabla u_j)$ and
$[\cdot]$ denotes the jump in stress across $\partial \Omega.$ 
\end{proposition}
As described in Specific Aim 3, we solve \eqref{SL1}--\eqref{SL3}  using a double layer potential representation.
While accurate in the far-field, this representation involves a strongly singular integral which
leads to large numerical errors when evaluating the the hydrophobic stress \eqref{stress} along the particle boundaries.
The identity in Proposition \ref{prop:recip} gets around this critical problem by utilizing the fact that $u_j$, $j\neq i$ is smooth and the normal derivative
of  $u_i$ continuous across $\partial P_i$.

Additionally, hydrodynamics interactions are also relevant the fabrication of complex microscopic three-dimensional structures \cite{Cho2010}.
Over the past decade, there has been an explosion of interest in small-scale processes that utilize capillary forces dominate,  van der Waals interactions and thermal noise  \cite{Zhang2017}, to coordinated movement and bind  material subunits. Two prominent fabrication techniques are capillary origami 
\cite{Pandey2011,Leong2007,Reynolds2019} and colloidal self-assembly \cite{Dasgupta2017,Siontorou2017}. Capillary origami uses principles of elastocapilirity wherein elastic solids deform under surface tension \cite{Bico2018,VanHonschoten2010}. In self-assembly of colloids, 
particles exhibit similar thermodynamic behavior of molecular systems, except that the behavior occurs over observable, long time scales \cite{Zhang2017}. 
and can determine which local equilibria the system tends towards.
